\documentclass[12pt]{article}
\usepackage[utf8]{inputenc}
\usepackage{geometry}
\geometry{a4paper, margin=1in}
\usepackage{tabularx}
\usepackage{booktabs}
\usepackage{pdflscape}
\usepackage{amsmath}
\usepackage{setspace} % Added for the onehalfspace environment to compile

\title{\vspace{4cm} \textbf{\Huge MoshiLite Distillation Report}}
\author{\Large S M Ridwan Husain Rafi -- 223 1349 642 \\
\Large Samin Al Azmain Ahmed -- 2311851042 \\
\Large Md. Nazmul Hasan Sakib -- 2312407042}
\date{}

\begin{document}

\begin{titlepage}
    \centering
    \vspace*{8cm}
    {\Huge S M Ridwan Husain Rafi -- 223 1349 642 \par}
    \vspace{0.5cm}
    {\Huge Samin Al Azmain Ahmed -- 2311851042 \par}
    \vspace{0.5cm}
    {\Huge Md. Nazmul Hasan Sakib -- 2312407042 \par}
    \vfill
\end{titlepage}

\newpage
\begin{landscape}
\section*{Training Summary \& Post-KD Results}

\begin{table}[h!]
\centering
\resizebox{\linewidth}{!}{
\begin{tabular}{lrrrr}
\toprule
\textbf{Model Variant} & \textbf{text} & \textbf{audio\_cb0} & \textbf{text\_top5} & \textbf{audio\_cb0} \\
 & \textbf{token\_acc} & \textbf{token\_acc} & \textbf{\_agree} & \textbf{\_top5\_agree} \\
\midrule
v1\_scattered\_nonuni\_L2 & 0.8249 & 0.2506 & 0.9088 & 0.5484 \\
v1\_scattered\_uni\_L2 & 0.8249 & 0.2506 & 0.9096 & 0.5484 \\
v2a\_cont\_strict\_nonuni\_L2 & 0.8249 & 0.2506 & 0.9102 & 0.5045 \\
v2a\_cont\_strict\_uni\_L2 & 0.8249 & 0.2534 & 0.9097 & 0.5662 \\
v2b\_cont\_penalized\_nonuni\_L2 & 0.8249 & 0.2506 & 0.9105 & 0.5484 \\
v2b\_cont\_penalized\_uni\_L2 & 0.8249 & 0.2506 & 0.9042 & 0.5611 \\
v2c\_cont\_relaxed\_nonuni\_L2 & 0.8249 & 0.2506 & 0.9088 & 0.5045 \\
v2c\_cont\_relaxed\_uni\_L2 & 0.8249 & 0.2507 & 0.9117 & 0.5687 \\
v3\_collapse\_nonuni\_L2 & 0.8249 & 0.2599 & 0.9046 & 0.5602 \\
v3\_collapse\_uni\_L2 & 0.8249 & 0.2507 & 0.9069 & 0.5666 \\
\midrule
magnitude\_30pct\_L2 & 0.9303 & 0.8242 & 0.9869 & 0.9725 \\
\bottomrule
\end{tabular}
}
\end{table}

\newpage
\subsection*{KL Divergence, Perplexity \& Validation Loss}
\begin{table}[h!]
\centering
\resizebox{\linewidth}{!}{
\begin{tabular}{lrrrrr}
\toprule
\textbf{Model Variant} & \textbf{text} & \textbf{audio\_cb0} & \textbf{text} & \textbf{audio\_cb0} & \textbf{val\_loss} \\
 & \textbf{kl\_div} & \textbf{kl\_div} & \textbf{perp.} & \textbf{perp.} & \textbf{\_l2} \\
\midrule
v1\_scattered\_nonuni\_L2 & 0.0007 & 0.0005 & 3.81 & 58.17 & 1.6215 \\
v1\_scattered\_uni\_L2 & 0.0007 & 0.0004 & 3.44 & 54.64 & 1.5720 \\
v2a\_cont\_strict\_nonuni\_L2 & 0.0008 & 0.0004 & 3.15 & 55.66 & 1.5509 \\
v2a\_cont\_strict\_uni\_L2 & 0.0006 & 0.0003 & 3.63 & 57.35 & 1.6018 \\
v2b\_cont\_penalized\_nonuni\_L2 & 0.0007 & 0.0004 & 3.13 & 59.73 & 1.5701 \\
v2b\_cont\_penalized\_uni\_L2 & 0.0005 & 0.0003 & 5.81 & 128.64 & 1.9855 \\
v2c\_cont\_relaxed\_nonuni\_L2 & 0.0139 & 0.0219 & 3.38 & 61.94 & 1.6280 \\
v2c\_cont\_relaxed\_uni\_L2 & 0.0005 & 0.0003 & 3.26 & 54.27 & 1.5533 \\
v3\_collapse\_nonuni\_L2 & 0.0021 & 0.0005 & 3.58 & 50.03 & 1.5582 \\
v3\_collapse\_uni\_L2 & 0.0006 & 0.0002 & 3.50 & 55.95 & 1.5837 \\
\midrule
magnitude\_30pct\_L2 & 0.0001 & 0.0006 & 1.30 & 1.83 & 0.2611 \\
\bottomrule
\end{tabular}
}
\end{table}

\newpage
\section*{Pre-KD vs Post-KD Distillation Comparison}

\begin{table}[h!]
\centering
\resizebox{\linewidth}{!}{
\begin{tabular}{l|rr|rr}
\toprule
\textbf{Model Variant} & \multicolumn{2}{p{3cm}|}{\centering\textbf{text\_top5}\\ \textbf{\_agree}} & \multicolumn{2}{p{3cm}}{\centering\textbf{audio\_cb0}\\ \textbf{\_top5\_agree}} \\
& Pre & Post & Pre & Post \\
\midrule
v1\_scattered\_nonuni\_L2 & 0.0053 & 0.9088 & 0.0032 & 0.5484 \\
v1\_scattered\_uni\_L2 & 0.0043 & 0.9096 & 0.0049 & 0.5484 \\
v2a\_cont\_strict\_nonuni\_L2 & 0.0061 & 0.9102 & 0.0036 & 0.5045 \\
v2a\_cont\_strict\_uni\_L2 & 0.0031 & 0.9097 & 0.0060 & 0.5662 \\
v2b\_cont\_penalized\_nonuni\_L2 & 0.0061 & 0.9105 & 0.0036 & 0.5484 \\
v2b\_cont\_penalized\_uni\_L2 & 0.0031 & 0.9042 & 0.0060 & 0.5611 \\
v2c\_cont\_relaxed\_nonuni\_L2 & 0.0061 & 0.9088 & 0.0036 & 0.5045 \\
v2c\_cont\_relaxed\_uni\_L2 & 0.0031 & 0.9117 & 0.0060 & 0.5687 \\
v3\_collapse\_nonuni\_L2 & 0.0075 & 0.9046 & 0.0040 & 0.5602 \\
v3\_collapse\_uni\_L2 & 0.0059 & 0.9069 & 0.0040 & 0.5666 \\
\midrule
magnitude\_30pct\_L2 & 0.9925 & 0.9869 & 0.9503 & 0.9725 \\
\bottomrule
\end{tabular}
}
\end{table}

\begin{table}[h!]
\centering
\resizebox{\linewidth}{!}{
\begin{tabular}{l|rr|rr}
\toprule
\textbf{Model Variant} & \multicolumn{2}{p{3cm}|}{\centering\textbf{text\_kl}\\ \textbf{\_div}} & \multicolumn{2}{p{3cm}}{\centering\textbf{audio\_cb0}\\ \textbf{\_kl\_div}} \\
& Pre & Post & Pre & Post \\
\midrule
v1\_scattered\_nonuni\_L2 & 0.0322 & 0.0007 & 0.0843 & 0.0005 \\
v1\_scattered\_uni\_L2 & 0.0226 & 0.0007 & 0.0821 & 0.0004 \\
v2a\_cont\_strict\_nonuni\_L2 & 0.0328 & 0.0008 & 0.0845 & 0.0004 \\
v2a\_cont\_strict\_uni\_L2 & 0.0231 & 0.0006 & 0.0831 & 0.0003 \\
v2b\_cont\_penalized\_nonuni\_L2 & 0.0328 & 0.0007 & 0.0845 & 0.0004 \\
v2b\_cont\_penalized\_uni\_L2 & 0.0231 & 0.0005 & 0.0831 & 0.0003 \\
v2c\_cont\_relaxed\_nonuni\_L2 & 0.0328 & 0.0139 & 0.0845 & 0.0219 \\
v2c\_cont\_relaxed\_uni\_L2 & 0.0231 & 0.0005 & 0.0831 & 0.0003 \\
v3\_collapse\_nonuni\_L2 & 0.0306 & 0.0021 & 0.0864 & 0.0005 \\
v3\_collapse\_uni\_L2 & 0.0228 & 0.0006 & 0.0832 & 0.0002 \\
\midrule
magnitude\_30pct\_L2 & 0.0068 & 0.0001 & 0.0869 & 0.0006 \\
\bottomrule
\end{tabular}
}
\end{table}
\newpage

\begin{table}[h!]
\centering
\resizebox{\linewidth}{!}{
\begin{tabular}{l|rr|rr|rr}
\toprule
\textbf{Model Variant} & \multicolumn{2}{p{2.5cm}|}{\centering\textbf{text}\\ \textbf{\_perplexity}} & \multicolumn{2}{p{2.5cm}|}{\centering\textbf{audio\_cb0}\\ \textbf{\_perplexity}} & \multicolumn{2}{p{2.5cm}}{\centering\textbf{val\_loss}\\ \textbf{\_l2}} \\
& Pre & Post & Pre & Post & Pre & Post \\
\midrule
v1\_scattered\_nonuni\_L2 & 152063.2 & 3.81 & 12793.4 & 58.17 & 6.4982 & 1.6215 \\
v1\_scattered\_uni\_L2 & 115073.1 & 3.44 & 14504.0 & 54.64 & 6.4439 & 1.5720 \\
v2a\_cont\_strict\_nonuni\_L2 & 5870.8 & 3.15 & 7514.0 & 55.66 & 5.3628 & 1.5509 \\
v2a\_cont\_strict\_uni\_L2 & 42785.6 & 3.63 & 41474.4 & 57.35 & 6.4634 & 1.6018 \\
v2b\_cont\_penalized\_nonuni\_L2 & 5870.8 & 3.13 & 7514.0 & 59.73 & 5.3628 & 1.5701 \\
v2b\_cont\_penalized\_uni\_L2 & 42785.6 & 5.81 & 41474.4 & 128.64 & 6.4634 & 1.9855 \\
v2c\_cont\_relaxed\_nonuni\_L2 & 5870.8 & 3.38 & 7514.0 & 61.94 & 5.3628 & 1.6280 \\
v2c\_cont\_relaxed\_uni\_L2 & 42785.6 & 3.26 & 41474.4 & 54.27 & 6.4634 & 1.5533 \\
v3\_collapse\_nonuni\_L2 & 64946.9 & 3.58 & 9395.0 & 50.03 & 6.1507 & 1.5582 \\
v3\_collapse\_uni\_L2 & 127345.1 & 3.50 & 244744.2 & 55.95 & 7.3230 & 1.5837 \\
\midrule
magnitude\_30pct\_L2 & 1.32 & 1.30 & 2.36 & 1.83 & 0.4054 & 0.2611 \\
\bottomrule
\end{tabular}
}
\end{table}
\end{landscape}

\newpage
\section{Dataset Collection}

The dataset for our knowledge distillation (KD) pipeline was synthetically generated via a self-play mechanism using a full-precision Moshi teacher model. 

Moshi can talk to itself via two channels, where a channel's output from the $t - 1$ timestamp is used as another channel's input at the $t$ timestamp.

The conversation sessions were initialized using random acoustic seeds comprising:
\begin{itemize}
    \item \textbf{Random noise vectors:} 80\%
    \item \textbf{Naturalistic conversational cues:} 10\%
    \item \textbf{Plain silence vectors:} 10\%
\end{itemize}

A 4-hour (375 steps at 12.5 Hz) audio corpus, containing 480 conversations of 30 seconds in length each, was generated. This process generally requires about 6--8 hours on an L4 GPU and 2--4 hours on an A100 GPU.

During generation, different filters were applied to prevent degenerated audio.

\subsection{Inner Monologue Filters}
\begin{itemize}
    \item \textbf{Length Check:} Rejects any conversation with fewer than 100 valid steps.
    \item \textbf{Meaningful Text Repetition:} Rejects if more than 50\% of the 4-grams on non-PAD tokens are repeated.
    \item \textbf{Second Half Repetition:} Addresses conversations that start normally but fall into an infinite loop in the second half. Rejects if more than 50\% of the 4-grams on non-PAD tokens from the final 15 seconds are repeated.
    \item \textbf{All-Token Text Repetition:} Prevents over-generation of identical data. Rejects if more than 92\% of all 4-grams are repeated. (This includes PAD tokens; the threshold is kept high since all silence/empty talks mathematically count as repetition).
\end{itemize}

\subsection{Audio/Codebook 0 (CB0) Filters}
We apply filters specifically to Codebook 0, as it contains the most important semantic and acoustic information.
\begin{itemize}
    \item \textbf{CB0 Repetition:} Removes acoustic loop generations. Rejects if more than 92\% of 4-grams in CB0 are repeated.
    \item \textbf{Silence Collapse:} Rejects if more than 80\% of the CB0 outputs are silence tokens.
\end{itemize}

\section{What Was Saved?}

\subsection{Hard Tokens (Exact decisions from the teacher model)}
\begin{itemize}
    \item \textbf{Channel A Audio Tokens:} This channel acted as the simulated ``user'' input to the teacher (Channel B). The audio produced by this channel is later used as input for the pruned student model.
    \item \textbf{Text Tokens (Inner Monologue):} The final chosen text tokens outputted by the model (\texttt{int16}). These tokens act as the core textual ground truth of what the teacher actually ``wanted'' to say. (Note: Generating these tokens increased audio accuracy by 3x).
    \item \textbf{Audio Tokens (8 Codebooks):} All 8 layers of audio RVQ codebook outputs.
\end{itemize}

\subsection{Sparse Logits (Probabilities behind decisions)}
\begin{itemize}
    \item \textbf{Text Logits (Top-50):} The teacher's probability distribution over the 32,000-word text vocabulary.
    \item \textbf{Audio CB0 Logits (Top-50):} The teacher's probability distribution over the 2,048-token Codebook 0 vocabulary. (The remaining codebooks were excluded).
\end{itemize}

\section{Training}

Moshi operates autoregressively. To predict the token for the $t^{th}$ timestamp, it feeds itself the user input plus all self-generated tokens up to the $t^{th}$ timestamp, and then predicts the token for the $(t + 1)^{th}$ timestamp. This process applies to both the inner monologue (text) and audio (Mimi RVQ).

In offline distillation, fully tracking this process is prone to errors. We bypassed this bottleneck by utilizing teacher forcing and Helium's masked self-attention mechanism (causal masking).

A 30-second user conversation (full Moshi Channel A outputs) is fed into the model. Due to causal masking, the student computes its prediction for the $T^{th}$ timestamp using only the user inputs and self-generated outputs from the $1, 2, 3, \dots, (T - 1)^{th}$ timestamps. Across all timestamps, predictions are calculated in parallel for all 375 steps.

\section{Loss Formulation}

\subsection{Step 1: The ``Soft'' Calculation (Sparse Logit KL-Divergence)}
For every timestep $t$, the student compares its probability outputs to the teacher's top-50 probabilities.

\begin{enumerate}
    \item \textbf{Softening the Probabilities:} The teacher's top-K sparse raw logits ($V^{(T)}$) and the student's raw logits extracted at those exact same top-K indices ($V^{(S)}$) are smoothed using a Temperature scalar ($\tau = 3.0$) via the Softmax function.
    \item \textbf{Measuring the Divergence:} The soft loss is calculated using KL-Divergence:
    \[
    \mathcal{L}_{logit} = D_{KL}({\mathrm{Teacher\_Logits}}_{t} \parallel {\mathrm{Student\_Logits}}_{t})
    \]
    \item \textbf{Aggregation:} This is calculated separately for the text vocabulary and the audio codebook vocabulary:
    \[
    \mathcal{L}_{soft\_total} = \mathcal{L}_{soft\_text} + \mathcal{L}_{soft\_audio}
    \]
\end{enumerate}

This aggregate makes up \textbf{70\%} (weight factor $\alpha = 0.7$) of the total penalty.

\subsection{Step 2: The ``Hard'' Calculation (Standard Cross-Entropy)}
Simultaneously, the student looks at the exact single token the teacher chose for timestep $t$.

Using the Cross-Entropy algorithm over the entire vocabulary space ($V$), it calculates the negative log-likelihood. The student is penalized purely based on whether it assigned the highest probability mass to the exact target token index chosen by the teacher:
\[
\mathcal{L}_{hard} = \text{CE}({\mathrm{Student\_Logits}}_{t}, {\mathrm{Teacher\_Chosen\_Token}}_{t})
\]

Like the soft loss, this is calculated for both text and audio tokens:
\[
\mathcal{L}_{hard\_total} = \mathcal{L}_{hard\_text} + \mathcal{L}_{hard\_audio}
\]

This operation is identical to standard large language model autoregressive training objectives. It makes up the remaining \textbf{30\%} (weight factor $\delta = 0.3$) of the total L2 penalty.

Each variant is trained for \textbf{3 epochs}.

\section{Evaluation Metrics}

To assess the pruned student models across multiple facets during and post-KD, a mimicking-based evaluation topology is utilized. This relies on quantitative divergence and categorical alignment mappings to estimate exact distributional fidelity:

\begin{itemize}
    \item \textbf{Text and Audio Token Accuracy (\texttt{text\_token\_acc}, \texttt{audio\_cb0\_token\_acc}):} Measures traditional token generation accuracy---whether the highest probability scalar prediction inferred by the student network directly matches the exact unigram chosen by the teacher.
    \begin{itemize}
        \item \texttt{argmax(student\_text\_logits) == teacher\_text\_token} (averaged over valid timesteps)
        \item \texttt{argmax(student\_audio\_logits) == teacher\_audio\_cb0\_token} (averaged over valid timesteps)
    \end{itemize}

    \item \textbf{Top-5 Agreement (\texttt{text\_top5\_agree}, \texttt{audio\_cb0\_top5\_agree}):} A relaxed intersection ratio measuring the overlap in the leading 5 vocabulary predictions. This dictates the student’s capability to retain the underlying contextual ambiguity within conversational audio flows as modeled by the teacher.
    \begin{itemize}
        \item Check if \texttt{teacher\_text\_token} $\in$ \texttt{top5(student\_text\_logits)} (averaged over valid timesteps)
        \item Check if \texttt{teacher\_audio\_cb0} $\in$ \texttt{top5(student\_audio\_logits)} (averaged over valid timesteps)
    \end{itemize}

    \item \textbf{KL Divergence (\texttt{text\_kl\_div}, \texttt{audio\_cb0\_kl\_div}):} Scales the distance spanning both logit spaces to pinpoint and statistically verify overall entropy shifts throughout predictions in text modeling paths over temporal domains.
    \begin{itemize}
        \item \texttt{KL(teacher\_text\_probs} $\parallel$ \texttt{student\_text\_probs)} using sparse top-50 logits, temperature-scaled ($\tau^2$), averaged over valid timesteps.
        \item \texttt{KL(teacher\_CB0\_probs} $\parallel$ \texttt{student\_CB0\_probs)} using sparse top-50 logits, temperature-scaled ($\tau^2$), averaged over valid timesteps.
    \end{itemize}

    \item \textbf{Perplexity (\texttt{text\_perplexity}, \texttt{audio\_cb0\_perplexity}):} Exponentiated entropy scores monitoring global predictive uncertainties. Drastic reductions seen in pre-KD and post-KD metrics act as fundamental verification logic demonstrating recovered linguistic representations.
    \begin{itemize}
        \item $\exp(\text{mean}(\text{CE}(\text{student\_text\_logits, teacher\_text\_token})))$ over valid timesteps. Capped at $\exp(20) \approx 485\text{M}$ to prevent overflow.
        \item $\exp(\text{mean}(\text{CE}(\text{student\_CB0\_logits, teacher\_CB0\_token})))$ over valid timesteps. Capped at $\exp(20) \approx 485\text{M}$ to prevent overflow.
    \end{itemize}

    \item \textbf{Validation Loss:} The primary comparative magnitude tracking structural proximity over offline target optimization. It scales directly with the KD training loss gradients to guide parametric optimizations stably.
    \begin{itemize}
        \item $\alpha \cdot \mathcal{L}_{\text{logit}}(\text{text} + \text{audio}) + \delta \cdot \mathcal{L}_{\text{hard}}(\text{text} + \text{audio})$ (with default weights $\alpha=0.7$, $\delta=0.3$)
    \end{itemize}
\end{itemize}

\begin{onehalfspace}
\section{Next To Do}
\begin{itemize}
    \item Since magnitude pruning generates the best outputs, we can different algorithms with different sparsity for it: global/iterative at 35%, 40%, 45%, 50%, 55%.

    \item 65\% of the layers were frozen during training of the structured pruned models. Therefore, more ca n be pruned from the versions. Test further sparsity.
    
    \item Use 2 moshi to generate better data for training/calibration instead of self-play. Each conversation should be at least 5 minutes long. Introduce filters to filter out degenerated/repeated/unusable audio.
    
    \item Explore other unstructured pruning algorithms (e.g., weight-based, magnitude-based, Wanda).
    \item \textbf{Top K:} Use top k = 500 or 1k instead of 50.
\end{itemize}
\end{onehalfspace}

\end{document}